---
title: "Advent Of Code"
author: "Issa Sanogo"
date: "03/12/2020"
output: html_document
---

```{r setup, include=FALSE}
# Options
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
options(stringsAsFactors = FALSE)

# Packages
library(tidyverse)
library(glue)
```

# AOC 2020

## --- Day 1 : Report Repair ---

```{r day1-input}
# read the input
input <- readLines("./data/2020/day1-input.txt")
```

### Part One

```{r part one}
two_entrie <- combn(input, 2) %>% # create a matrix of possible
  as.tibble() %>%
  map_df(as.numeric)
two_entrie <- unlist(two_entrie[colSums(two_entrie) == 2020], use.names = FALSE)
# two_entrie
# sum(two_entrie)
# prod(two_entrie)
```

In my list, the two entries that sum to `2020` are `r paste(two_entrie, collapse = " and ")`.
Multiplying them together produces `r paste(two_entrie, collapse = " * ")` = `r as.character(prod(two_entrie))`.

### Part Two

```{r part two}
three_entrie <- combn(input, 3) %>%
  as.tibble() %>%
  map_df(as.numeric)
three_entrie <- unlist(three_entrie[colSums(three_entrie) == 2020], use.names = FALSE)
# three_entrie
# sum(three_entrie)
# prod(three_entrie)
```

In my list, the three entries that sum to `2020` are `r glue_collapse(three_entrie, ", ", last = ", and ")`. Multiplying them together produces `r paste(three_entrie,  collapse = " * ")` = `r as.character(prod(three_entrie))`.
